IJAR Bayesian Nonparametrics Special Issue Manuscript

### 2015-10-20 ijar-npb-ms.tex
minor changes sec1-3

@@

Based on expert assumptions and data for component failure distribution,
we calculate the system reliability function $\Rsys(t) = P(\Tsys > t)$.

@@

Alternative to system signature that allows for
straightforward computation of system reliability
for multiple component types.
It separates the (time-invariant) system structure from the time-dependent failure probabilities of components.

@@

The random failure time of component number $i$ of type $k$ is denoted by $T^k_i$, $i = 1, \ldots m_k$.
For its distribution, written in terms of the cdf $F^k(t) = P(T^k_i \le t)$
or in terms of the reliability function $R^k(t) = P(T^k > t) = 1 - F^k(t)$,
we consider a set of time points $t$.

For each time point $t$, the functioning of a single component of type $k$ at time $t$ (functioning: 1, failed: 0)
is Bernoulli distributed with a suitable functioning probability (as, $1-$ failure probability) $p^k_t$,
$\indic(T^k_i > t) \sim \ber(p^k_t)$, $i = 1, \ldots, m_k$.

Then, the number of functioning components out of the $m_k$ components of type $k$
is binomially distributed, $C^k_t = \sum_{i=1}^{m_k} \indic(T^k_i > t) \sim \bin(p^k_t, m_k)$.

@@

***or say we don't know for times between time points, or lower and upper bound already here?***

(maybe not relevant for this manuscript, but would it make sense to take Kaplan-Meier estimates for the $p^k_t$'s,
or the NPI bounds for Kaplan-Meier?)

***Right-censored observations maybe in a second step?
Minimal assumption (component can fail immediately after censoring or live forever)
will lead to high imprecision,
assumption of exchangeability with other components at moment of censoring
will be quite complex, maybe too much for this paper,
and could be covered in a second paper with extensive use-case?***

The $p^k_t$'s can be chosen such that they reflect, e.g., a bathtub curve for the corresponding hazard rate.

***Constraint: $p^k_{t_j} \ge p^k_{t_{j+1}}$ must hold (assuming no repair).***

@@

Having observed the lifetime data $\vec{t}^k = (t^k_1, \ldots, t^k_{n_k})$,
at any fixed time $t$ this results in the observation from the Binomial model described above,
$s^k_t = \sum_{i=1}^{n_k} \indic(t^k_i > t)$.
The posterior is then $p^k_t \mid s^k_t \sim \be(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
The combination of a Binomial observation model with a Beta prior is often called Beta-binomial model.

@@

Let us drop the super- and subscript $k$ and $t$ for a while
to discuss the role and interpretation of the parameters $\alpha$ and $\beta$
for each $k$ and $t$.
The Beta distribution is usually parametrized in terms of $\alpha$ and $\beta$ as used above, 
but for our generalization to sets of priors,
it is useful to consider a different parametrization in terms of $\nz$, $\yz$:
\begin{align*}
\nz &= \alpha + \beta &
\yz &= \frac{\alpha}{\alpha + \beta} \,,
\end{align*}
or vice versa, $\alpha = \nz\yz$ and $\beta = \nz(1-\yz)$.
From the properties of the Beta distribution,
we see that $\yz = \E[p]$ is the prior mean for the functioning probability $p$,
and that the higher $\nz$, the more probability weight will be concentrated around that mean,
as $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Furthermore, $\nz$ can be interpreted as a pseudocount or prior strength,
this will be illustrated through revisiting the posterior
based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$:

$p \mid s$ is Beta distributed with the updated parameters
\begin{align*}
\nn &= \nz + n &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,,
\end{align*}
so after seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of function components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Also, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

@@

\section{Conclusions}

>>>>

(some sentences may be useful for the abstract)

@@

Based on expert assumptions and data for component failure distribution,
the goal is to calculate the system reliability function $\Rsys(t) = P(\Tsys > t)$.
Considering a discrete grid of time points $\{t_1, \ldots, \tmax\}$,
naturally a Bayesian model for the functioning probability at each time point arises.
To account for uncertainty in the choice of prior and due to the inherent issue with prior-data conflict,
we propose an impürecise or interval probability approach
based on sets of prior distributions defined through sets of canonical parameters.

Our method produces a set of discrete system reliability functions,
or, for each time point $t$, an interval for the system survival probability
that appropriately reflects uncertainty due to vague prior information, the amount of data, and prior-data conflict.

@@

The survival signature allows for straightforward computation of system reliability
for systems with complex layouts and multiple component types.
It separates the (time-invariant) system structure from the time-dependent failure probabilities of components.

@@

The random failure time of component number $i$ of type $k$ is denoted by $T^k_i$, $i = 1, \ldots m_k$.
For its distribution, written in terms of the cdf $F^k(t) = P(T^k_i \le t)$
or in terms of the reliability function $R^k(t) = P(T^k > t) = 1 - F^k(t)$,
we consider a set of time points $t$, $t \in {\cal T} = \{t_1, \ldots, \tmax\}$.

For each time point $t$, the functioning of a single component of type $k$ at time $t$ (functioning: 1, failed: 0)
is Bernoulli distributed with a suitable functioning probability (as, $1-$failure probability) $p^k_t$, so
\begin{align*}
P\big(\indic(T^k_i > t) = 1\big) &= p^k_t\,, \\
P\big(\indic(T^k_i > t) = 0\big) &= 1 - p^k_t\,,
\end{align*}
in short, $\indic(T^k_i > t) \sim \ber(p^k_t)$, $i = 1, \ldots, m_k$.

Then, due to the assumption of independence of components of the same type, 
the number of functioning components out of the $m_k$ components of type $k$
is binomially distributed, $C^k_t = \sum_{i=1}^{m_k} \indic(T^k_i > t) \sim \bin(p^k_t, m_k)$.

@@

***or say we don't know for times between time points, or lower and upper bound already here?***
***illustrate with a graph?***

%(maybe not relevant for this manuscript, but would it make sense to take Kaplan-Meier estimates for the $p^k_t$'s,
%or the NPI bounds for Kaplan-Meier?)
%

The $p^k_t$'s can be taken such that they reflect, e.g., a bathtub curve for the corresponding hazard rate.
Naturally, $p^k_{t_j} \ge p^k_{t_{j+1}}$ should hold (assuming no repair).***

@@

Having observed the lifetime data $\vec{t}^k = (t^k_1, \ldots, t^k_{n_k})$,
at any fixed time $t \in {\cal T}$ this results in the observation from the Binomial model described above,
$s^k_t = \sum_{i=1}^{n_k} \indic(t^k_i > t)$.
The posterior is then $p^k_t \mid s^k_t \sim \be(\alpha^k_t + s^k_t, \beta^k_t + n_k - s^k_t)$.
The combination of a Binomial observation model with a Beta prior is often called Beta-binomial model.

@@

Let us drop the super- and subscript $k$ and $t$ for a while
to discuss the role and interpretation of the parameters $\alpha$ and $\beta$
for each $k$ and $t$.
The Beta distribution is usually parametrized in terms of $\alpha$ and $\beta$ as used above, 
but for our generalization to sets of priors,
it is useful to consider a different parametrization in terms of $\nz$ and $\yz$, where
\begin{align*}
\nz &= \alpha + \beta &
&\text{and} &
\yz &= \frac{\alpha}{\alpha + \beta} \,,
\end{align*}
or vice versa, $\alpha = \nz\yz$ and $\beta = \nz(1-\yz)$.
$\nz$ and $\yz$ are sometimes known as the \emph{canonical} parameters,
identified from rewriting the density in a canonical form.
(This canonical form gives a joint structure to all conjugate priors.)

From the properties of the Beta distribution,
we see that $\yz = \E[p]$ is the prior expectation for the functioning probability $p$,
and that the higher $\nz$, the more probability weight will be concentrated around that mean,
as $\V(p) = \frac{\yz (1-\yz)}{\nz + 1}$.
Furthermore, $\nz$ can be interpreted as a pseudocount or prior strength,
this will be illustrated through revisiting the posterior
based on observed lifetimes $\vec{t} = (t_1, \ldots, t_{n})$:

$p \mid s$ is Beta distributed with the updated parameters
\begin{align*}
\nn &= \nz + n\,, &
\yn &= \frac{\nz}{\nz + n} \cdot \yz + \frac{n}{\nz + n} \cdot \frac{s}{n}\,,
\end{align*}
so after seeing that $s$ out of $n$ components function (at time $t$),
the posterior mean $\yn$ for $p$ is a weighted average of
the prior mean $\yz$ and $s/n$ (the fraction of function components in the data),
with the weights $\nz$ and $n$, respectively.
We see that $\nz$ plays the same role for the prior mean $\yz$
as the sample size $n$ for the observed mean $s/n$,
thus the notion of pseudocount.
Also, the higher $\nz$, the higher the weight for $\yz$
in the weighted average calculation of $\yn$,
so $\nz$ gives the strength of the prior as compared to the sample size $n$.
%The parameters $\nz$ and $\yz$ have a more intuitive interprtation,

@@

Also calculate $\Rsys(t)$ intervals for vacuous prior ($[\yktzl, \yktzu] = [0,1]$, some fixed $\nktz$)?***

@@

\section{Conclusions and Outlook}

Extend the model to deal with right-censored observations which are common in realibility setting.
We assume that a minimal assumption (component can fail immediately after censoring or live forever)
will be simple to implememt but will lead to high imprecision,
whereas assuming exchangeability with other components at moment of censoring
will be more complex to accomodate but will lead to les imprecision.
